---
title: "Probability of covid-19 infection given RT-PCR negative"
author: "Stephen A Lauer"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
knitr::opts_knit$set(root.dir = here::here())

options(mc.cores=4,
        scipen=999)
```

```{r library}
library(tidyverse)
library(rstan)
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
n_iter <- 1e4
n_warmup <- 1e4-5e3
p_adapt_delta <- 0.99
n_max_treedepth <- 20
T_max <- 21

source("R/utils.R")

pcr_dat <- read_csv("data/antibody-test-data.csv") %>% 
    filter(grepl("RT_PCR", test), n > 0) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor()
           %>% as.numeric(),
           pct_pos=test_pos/n)

elisa_dat <- read_csv("data/antibody-test-data.csv") %>% 
    filter(!grepl("RT_PCR", test), n > 0) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor()
           %>% as.numeric(),
           pct_pos=test_pos/n)

igg_dat <- read_csv("data/antibody-test-data.csv") %>% 
    filter(test=="IgG", n > 0) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor()
           %>% as.numeric(),
           pct_pos=test_pos/n)

igm_dat <- read_csv("data/antibody-test-data.csv") %>% 
    filter(test=="IgM", n > 0) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor()
           %>% as.numeric(),
           pct_pos=test_pos/n)
```

In this paper, we will try to determine the probability that someone has covid-19 given a negative RT-PCR test.

## Methods

[Zhao et al. (2020)](https://academic.oup.com/cid/advance-article/doi/10.1093/cid/ciaa344/5812996), [Liu et al. (2020)](https://www.medrxiv.org/content/10.1101/2020.03.06.20031856v1), [Guo et al. (2020)](https://academic.oup.com/cid/article-abstract/doi/10.1093/cid/ciaa310/5810754), and [Wölfel et al. (2020)](https://www.nature.com/articles/s41586-020-2196-x) looked at the sensitivity of the RT-PCR (and ELISAs) by time since symptom onset.

```{r raw-figures}
ggplot(data=pcr_dat, aes(x=day, y=pct_pos, size=n, color=study)) +
    facet_grid(.~test) +
    geom_point(alpha=0.4) +
    scale_x_continuous("Days since symptom onset",
                       breaks=seq(0,35,7)) +
    scale_y_continuous("Sensitivity of the RT-PCR test") +
    scale_size_continuous("Number of\nobservations") +
    scale_color_manual("Study",
                       values=cbbPalette[-1]) +
    guides(colour = guide_legend(override.aes = list(alpha = 1))) +
    theme_bw()

ggplot(data=elisa_dat, aes(x=day, y=pct_pos, size=n, color=study)) +
    facet_grid(.~test) +
    geom_point(alpha=0.4) +
    scale_x_continuous("Days since symptom onset",
                       breaks=seq(0,35,7)) +
    scale_y_continuous("Sensitivity of the antibody test") +
    scale_size_continuous("Number of\nobservations") +
    scale_color_manual("Study",
                       values=cbbPalette[-1]) +
    guides(colour = guide_legend(override.aes = list(alpha = 1))) +
    theme_bw()
```

The sensitivity rises to a peak 4 days after symptom onset then declines for the next couple of weeks.

If we know the risk of an individual, we can find the negative predictive value -- the probability that someone who tests negative is actually negative.
From [Bi et al.](https://www.medrxiv.org/content/10.1101/2020.03.03.20028423v3), we know that about 15% (77/517) household contacts later tested positive for covid-19.

We use logistic regression for the sensitivity of the RT-PCR with a cubic polynomial for the log of time since exposure and use that, along with the probability of infection given exposure, to estimate the negative predictive value of the RT-PCR.
We use estimates of the incubation period from [Lauer, Grantz, et al. (2020)](https://annals.org/aim/fullarticle/2762808/incubation-period-coronavirus-disease-2019-covid-19-from-publicly-reported).
From this, we can find the probability of having a covid-19 infection despite being RT-PCR negative.

We use Stan for this analysis.

## Results

```{r npv-fixed-onset, cache=T, eval=T, include=F}
## fit a model to find the overall seroincidence across all observations
npv_onset_model <- stan_model("Stan/npv-fixed-onset.stan")

main_analysis <- make_analysis_data(stan_model=npv_onset_model,
                                    data_list=list(N=nrow(pcr_dat),
                                                   J=max(pcr_dat$study_idx),
                                                   T_max=T_max,
                                                   test_n=pcr_dat$n,
                                                   test_pos=pcr_dat$test_pos,
                                                   study_idx=pcr_dat$study_idx,
                                                   t_symp_test=pcr_dat$day,
                                                   exposed_n=686,
                                                   exposed_pos=77,
                                                   t_exp_symp=5,
                                                   spec=1),
                                    iter=n_iter,
                                    warmup=n_warmup,
                                    control=list(adapt_delta=p_adapt_delta,
                                                 max_treedepth=n_max_treedepth),
                                    save_warmup=F)
```

```{r fnr-figure}
fnr_fig <- ggplot(data=main_analysis$plot_dat, aes(x=days_since_exposure)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_errorbar(aes(ymin=fnr_lb, ymax=fnr_ub), color="gray30") +
    geom_point(aes(y=fnr_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability RT-PCR negative,\ngiven covid-19 positive",
                       limits=c(0,1)) +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())
```

With no data on RT-PCR to time prior to symptom onset, the estimates of sensitivity at or below day five are low with large credible intervals.

```{r npv-figure}
for_fig <- ggplot(data=main_analysis$plot_dat, aes(x=days_since_exposure)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_errorbar(aes(ymax=for_lb, ymin=for_ub), color="gray30") +
    geom_point(aes(y=for_med)) +
    scale_x_continuous("Days since exposure",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability infected,\ngiven RT-PCR negative",
                       limits=c(0, 0.15),
                       breaks=seq(0,0.15, 0.05)) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))

gridExtra::grid.arrange(fnr_fig, for_fig)

# ggplot(plot_dat, aes(x=days_since_exposure)) +
#     geom_hline(aes(yintercept=1), linetype="dashed") +
#     geom_errorbar(aes(ymax=1-rr_lb, ymin=1-rr_ub), color="gray30") +
#     geom_point(aes(y=1-rr_med)) +
#     scale_x_continuous("Days since exposure",
#                        breaks=seq(0, 21, 7),
#                        limits=c(0,21.5)) +
#     scale_y_log10("Relative risk of having a negative RT-PCR test",
#                   breaks=2^c(-3:1),
#                   labels=c("1/8", "1/4", "1/2", "1", "2")) +
#     coord_cartesian(ylim=c(2^-3, 2^1)) +
#     theme_bw() +
#     theme(axis.text=element_text(color="black"))
```


Due to the decline in sensitivity over time, the RT-PCR test is best deployed about a week after exposure.
A day or two after exposure (3 or 4 days prior to symptoms), the test may have no utility at all, and thus the probability of having been infected is would be the same with or without an RT-PCR, in our case about 11%.
Seven to nine days after exposure (roughly 2 to 4 days after symptom onset), the negative predictive value is around 95%, meaning there is about a 5% chance of actually being covid-19 positive despite testing negative.

\pagebreak

```{r npv-antibodies, cache=T, eval=T, include=F}
## fit a model to find the overall seroincidence across all observations
npv_antibodies_model <- stan_model("Stan/npv-antibodies.stan")

igg_analysis <- make_analysis_data(stan_model=npv_antibodies_model,
                                   data_list=list(N=nrow(igg_dat),
                                                  J=max(igg_dat$study_idx),
                                                  T_max=T_max,
                                                  test_n=igg_dat$n,
                                                  test_pos=igg_dat$test_pos,
                                                  study_idx=igg_dat$study_idx,
                                                  t_symp_test=igg_dat$day,
                                                  exposed_n=686,
                                                  exposed_pos=77,
                                                  t_exp_symp=5,
                                                  spec=1),
                                   iter=n_iter,
                                   warmup=n_warmup,
                                   control=list(adapt_delta=p_adapt_delta,
                                                max_treedepth=n_max_treedepth),
                                   save_warmup=F)


igm_analysis <- make_analysis_data(stan_model=npv_antibodies_model,
                                   data_list=list(N=nrow(igm_dat),
                                                  J=max(igm_dat$study_idx),
                                                  T_max=T_max,
                                                  test_n=igm_dat$n,
                                                  test_pos=igm_dat$test_pos,
                                                  study_idx=igm_dat$study_idx,
                                                  t_symp_test=igm_dat$day,
                                                  exposed_n=686,
                                                  exposed_pos=77,
                                                  t_exp_symp=5,
                                                  spec=1),
                                   iter=n_iter,
                                   warmup=n_warmup,
                                   control=list(adapt_delta=p_adapt_delta,
                                                max_treedepth=n_max_treedepth),
                                   save_warmup=F)
```

```{r antibody-fig, eval=F}
bind_rows(igg_analysis$plot_dat %>% mutate(test="IgG"),
          igm_analysis$plot_dat %>% mutate(test="IgM"),
          main_analysis$plot_dat %>% mutate(test="RT-PCR")) %>% 
    ggplot(aes(x=days_since_exposure, color=test)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    # geom_errorbar(aes(ymax=for_lb, ymin=for_ub), color="gray30") +
    geom_point(aes(y=for_med)) +
    scale_x_continuous("Days since exposure",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability infected,\ngiven RT-PCR negative",
                       limits=c(0, 0.15),
                       breaks=seq(0,0.15, 0.05)) +
    scale_color_manual("Test",
                       values=cbbPalette[c(2,3,1)]) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```
<!-- \pagebreak -->

## Sensitivity analyses

### Specificity

What if the specificity of the test is less than 100%?
To test this we fit the same model except with a specificity of 90% and compared the results to the original (the sensitivity remained the same).

```{r sens-analysis-spec, cache=T, eval=T}
spec_est <- make_analysis_data(stan_model=npv_onset_model,
                                    data_list=list(N=nrow(pcr_dat),
                                                   J=max(pcr_dat$study_idx),
                                                   T_max=T_max,
                                                   test_n=pcr_dat$n,
                                                   test_pos=pcr_dat$test_pos,
                                                   study_idx=pcr_dat$study_idx,
                                                   t_symp_test=pcr_dat$day,
                                                   exposed_n=686,
                                                   exposed_pos=77,
                                                   t_exp_symp=5,
                                                   spec=0.9),
                                    iter=n_iter,
                                    warmup=n_warmup,
                                    control=list(adapt_delta=p_adapt_delta,
                                                 max_treedepth=n_max_treedepth),
                                    save_warmup=F)
```

```{r plot-spec}
spec_est$plot_dat %>% 
    mutate(spec="90%") %>% 
    bind_rows(main_analysis$plot_dat %>% mutate(spec="100%")) %>% 
    ggplot(aes(x=days_since_exposure, color=spec, fill=spec)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_ribbon(aes(ymax=for_lb, ymin=for_ub), alpha=0.5) +
    # geom_errorbar(aes(ymax=for_lb, ymin=for_ub), alpha=0.5) +
    geom_line(aes(y=for_med)) +
    # geom_point(aes(y=for_med)) +
    scale_x_continuous("Days since exposure",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability infected,\ngiven RT-PCR negative",
                       limits=c(0, 0.16),
                       breaks=seq(0,0.15, 0.05)) +
    scale_color_manual("Specificity",
                       values=cbbPalette) +
    scale_fill_manual("Specificity",
                      values=cbbPalette) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```

The shape of the curve for 90% specificity is similar to that of 100%, though slightly elevated.
The best time to test is still 2-4 days post-symptom onset.

\pagebreak

### Different attack rates

We tested attack rates of half, twice, and four times that of Bi et al. (2020).

```{r ar-est, cache=T, eval=T}
half_est <- make_analysis_data(stan_model=npv_onset_model,
                                    data_list=list(N=nrow(pcr_dat),
                                                   J=max(pcr_dat$study_idx),
                                                   T_max=T_max,
                                                   test_n=pcr_dat$n,
                                                   test_pos=pcr_dat$test_pos,
                                                   study_idx=pcr_dat$study_idx,
                                                   t_symp_test=pcr_dat$day,
                                                   exposed_n=686,
                                                   exposed_pos=round(77/2),
                                                   t_exp_symp=5,
                                                   spec=1),
                                    iter=n_iter,
                                    warmup=n_warmup,
                                    control=list(adapt_delta=p_adapt_delta,
                                                 max_treedepth=n_max_treedepth),
                                    save_warmup=F)

two_est <- make_analysis_data(stan_model=npv_onset_model,
                              data_list=list(N=nrow(pcr_dat),
                                             J=max(pcr_dat$study_idx),
                                             T_max=T_max,
                                             test_n=pcr_dat$n,
                                             test_pos=pcr_dat$test_pos,
                                             study_idx=pcr_dat$study_idx,
                                             t_symp_test=pcr_dat$day,
                                             exposed_n=686,
                                             exposed_pos=2*77,
                                             t_exp_symp=5,
                                             spec=1),
                              iter=n_iter,
                              warmup=n_warmup,
                              control=list(adapt_delta=p_adapt_delta,
                                           max_treedepth=n_max_treedepth),
                              save_warmup=F)

four_est <- make_analysis_data(stan_model=npv_onset_model,
                               data_list=list(N=nrow(pcr_dat),
                                              J=max(pcr_dat$study_idx),
                                              T_max=T_max,
                                              test_n=pcr_dat$n,
                                              test_pos=pcr_dat$test_pos,
                                              study_idx=pcr_dat$study_idx,
                                              t_symp_test=pcr_dat$day,
                                              exposed_n=686,
                                              exposed_pos=4*77,
                                              t_exp_symp=5,
                                              spec=1),
                               iter=n_iter,
                               warmup=n_warmup,
                               control=list(adapt_delta=p_adapt_delta,
                                            max_treedepth=n_max_treedepth),
                               save_warmup=F)
```

```{r plot-ar}
bind_rows(half_est$plot_dat %>% mutate(ar_idx="half"),
          two_est$plot_dat %>% mutate(ar_idx="2x"),
          four_est$plot_dat %>% mutate(ar_idx="4x"),
          main_analysis$plot_dat %>% mutate(ar_idx="Bi")) %>% 
    ggplot(aes(x=days_since_exposure, fill=as.factor(ar_idx),
               color=as.factor(ar_idx))) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    # geom_ribbon(aes(ymin=for_lb, ymax=for_ub), alpha=0.3) +
    geom_line(aes(y=for_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability infected,\ngiven RT-PCR negative") +
    scale_color_manual("Attack rate",
                       values=cbbPalette[c(2,3,1,4)],
                       breaks=c("half", "Bi", "2x", "4x"),
                       labels=c("5.5%", "11%", "22%", "44%")) +
    scale_fill_manual("Attack rate",
                       values=cbbPalette[c(2,3,1,4)],
                       breaks=c("half", "Bi", "2x", "4x"),
                       labels=c("5.5%", "11%", "22%", "44%")) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```

As the attack rate increases, the shape looks more like the false negative rate seen above in Figure 2.
As the attack rate decreases, the probabilities approach 0 and thus the entire curve flattens out.
Regardless, the best time to test is still 2-4 days after symptom onset.

\pagebreak

### Different incubation period lengths

We originally assumed a 5-day incubation period, what if that was 3 or 7 days instead?

```{r inc-est, cache=T, eval=T}
three_day_est <- make_analysis_data(stan_model=npv_onset_model,
                                    data_list=list(N=nrow(pcr_dat),
                                                   J=max(pcr_dat$study_idx),
                                                   T_max=T_max,
                                                   test_n=pcr_dat$n,
                                                   test_pos=pcr_dat$test_pos,
                                                   study_idx=pcr_dat$study_idx,
                                                   t_symp_test=pcr_dat$day,
                                                   exposed_n=686,
                                                   exposed_pos=77,
                                                   t_exp_symp=3,
                                                   spec=1),
                                    iter=n_iter,
                                    warmup=n_warmup,
                                    control=list(adapt_delta=p_adapt_delta,
                                                 max_treedepth=n_max_treedepth),
                                    save_warmup=F)

seven_day_est <- make_analysis_data(stan_model=npv_onset_model,
                               data_list=list(N=nrow(pcr_dat),
                                              J=max(pcr_dat$study_idx),
                                              T_max=T_max,
                                              test_n=pcr_dat$n,
                                              test_pos=pcr_dat$test_pos,
                                              study_idx=pcr_dat$study_idx,
                                              t_symp_test=pcr_dat$day,
                                              exposed_n=686,
                                              exposed_pos=77,
                                              t_exp_symp=7,
                                              spec=1),
                               iter=n_iter,
                               warmup=n_warmup,
                               control=list(adapt_delta=p_adapt_delta,
                                            max_treedepth=n_max_treedepth),
                               save_warmup=F)
```

```{r plot-inc}
bind_rows(three_day_est$plot_dat %>% mutate(inc_period="3d"),
          seven_day_est$plot_dat %>% mutate(inc_period="7d"),
          main_analysis$plot_dat %>% mutate(inc_period="5d")) %>% 
    ggplot(aes(x=days_since_exposure, fill=as.factor(inc_period),
               color=as.factor(inc_period))) +
    geom_vline(aes(xintercept=3), linetype="dashed", color=cbbPalette[2]) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_vline(aes(xintercept=7), linetype="dashed", color=cbbPalette[3]) +
    # geom_ribbon(aes(ymin=for_lb, ymax=for_ub), alpha=0.1) +
    geom_line(aes(y=for_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability infected,\ngiven RT-PCR negative",
                       limits=c(0,0.15)) +
    scale_color_manual("Incubation period",
                       values=cbbPalette[c(2,1,3)]) +
    scale_fill_manual("Incubation period",
                      values=cbbPalette[c(2,1,3)]) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```

Changing the length of the incubation period changes the progression of the false omission rate (probability infected given test negative).
Since the sensitivity is calibrated with respect to the time of symptom onset, an earlier onset time leads to a quicker drop in false omission rate and a later onset time leads to a slower drop.

<!-- \pagebreak -->

<!-- ### Incubation period as a random variable -->

<!-- Since we don't know the exact incubation period time, what if we used a random variable for the incubation period based on [Lauer, Grantz, et al. (2020)](https://annals.org/aim/fullarticle/2762808/incubation-period-coronavirus-disease-2019-covid-19-from-publicly-reported)? -->

<!-- ```{r npv-rv, cache=T, eval=T, include=FALSE} -->
<!-- ## fit a model to find the negative predictive value using a random variable -->
<!-- ## for the incubation period -->
<!-- npv_rv_model <- stan_model("Stan/npv-analysis.stan") -->

<!-- npv_rv_est <- sampling(npv_rv_model, -->
<!--                        data=list(N=nrow(pcr_dat), -->
<!--                                  J=max(pcr_dat$study_idx), -->
<!--                                  T_max=T_max, -->
<!--                                  test_n=pcr_dat$n, -->
<!--                                  test_pos=pcr_dat$test_pos, -->
<!--                                  study_idx=pcr_dat$study_idx, -->
<!--                                  t_symp_test=pcr_dat$day, -->
<!--                                  exposed_n=686, -->
<!--                                  exposed_pos=77, -->
<!--                                  spec=1 -->
<!--                        ), -->
<!--                        iter=n_iter, -->
<!--                        warmup=n_warmup, -->
<!--                        control=list(adapt_delta=p_adapt_delta, -->
<!--                                     max_treedepth=n_max_treedepth), -->
<!--                        chains=6, -->
<!--                        save_warmup=F) -->
<!-- ``` -->

<!-- ```{r rv-conditionals} -->
<!-- # loo::extract_log_lik(npv_est) %>% loo::loo() -->
<!-- ## sensitivity (sens) of PCR: P(PCR+ | covid+) -->
<!-- ## false negative rate (fnr) of PCR: P(PCR- | covid+) = 1 - sens -->
<!-- sens_rv <- extract(npv_rv_est, pars="sens")[[1]] -->

<!-- ## negative predictive value (npv) of PCR: P(covid- | PCR-) -->
<!-- ## false omission rate (FOR) of PCR: P(covid+ | PCR-) = 1 - npv -->
<!-- npv_rv <- extract(npv_rv_est, pars="npv")[[1]] -->

<!-- ## attack rate: P(covid+) -->
<!-- ## P(covid-) = 1 - attack_rate -->
<!-- ar_rv <- extract(npv_rv_est, pars="attack_rate")[[1]] %>% as.vector() -->
<!-- ``` -->

<!-- ```{r plot-rv} -->
<!-- bind_rows(as_tibble(sens_rv) %>% -->
<!--               gather("days", "sens") %>% -->
<!--               mutate(days_since_exposure=gsub(pattern="V", "", days) %>% -->
<!--                          as.numeric) %>% -->
<!--               bind_cols(as_tibble(npv_rv) %>% -->
<!--                             gather("days", "npv") %>% -->
<!--                             mutate(ar=rep(ar_rv, T_max)) %>% -->
<!--                             select(-days)) %>% -->
<!--               group_by(days_since_exposure) %>% -->
<!--               summarise(fnr_med=median(1-sens), -->
<!--                         fnr_lb=quantile(1-sens,probs=.025), -->
<!--                         fnr_ub=quantile(1-sens,probs=.975), -->
<!--                         for_med=median(1-npv), -->
<!--                         for_lb=quantile(1-npv,probs=.025), -->
<!--                         for_ub=quantile(1-npv,probs=.975), -->
<!--                         rr_med=median(1-(1-npv)/ar), -->
<!--                         rr_lb=quantile(1-(1-npv)/ar,probs=.025), -->
<!--                         rr_ub=quantile(1-(1-npv)/ar,probs=.975), -->
<!--                         abs_med=median(ar-(1-npv)), -->
<!--                         abs_lb=quantile(ar-(1-npv),probs=.025), -->
<!--                         abs_ub=quantile(ar-(1-npv),probs=.975)) %>%  -->
<!--               mutate(type="rv"), -->
<!--           plot_dat %>% mutate(type="fixed")) %>%  -->
<!--     ggplot(aes(x=days_since_exposure, color=type, fill=type)) + -->
<!--     geom_vline(aes(xintercept=5), linetype="dashed") + -->
<!--     geom_ribbon(aes(ymax=for_lb, ymin=for_ub), alpha=0.4) + -->
<!--     # geom_errorbar(aes(ymax=for_lb, ymin=for_ub), color="gray30") + -->
<!--     geom_line(aes(y=for_med)) + -->
<!--     # geom_point(aes(y=for_med)) + -->
<!--     scale_x_continuous("Days since exposure", -->
<!--                        breaks=seq(0, 21, 7), -->
<!--                        limits=c(0,21.5)) + -->
<!--     scale_y_continuous("Probability infected,\ngiven RT-PCR negative", -->
<!--                        limits=c(0, 0.15), -->
<!--                        breaks=seq(0,0.15, 0.05)) + -->
<!--     scale_color_manual("Type", -->
<!--                        labels=c("Fixed onset", "Random variable"), -->
<!--                        values=cbbPalette) + -->
<!--     scale_fill_manual("Type", -->
<!--                       labels=c("Fixed onset", "Random variable"), -->
<!--                       values=cbbPalette) + -->
<!--     theme_bw() + -->
<!--     theme(axis.text=element_text(color="black")) -->
<!-- ``` -->

<!-- Treating the incubation period as a random variable adds uncertainty to the first week after exposure. -->
