---
title: "Probability of covid-19 infection given RT-PCR negative"
author: "Stephen A Lauer"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
knitr::opts_knit$set(root.dir = here::here())

options(mc.cores=4,
        scipen=999)
```

```{r library}
library(tidyverse)
library(rstan)
## color blind palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
## Stan settings
n_iter <- 1e4
n_warmup <- 1e4-5e3
p_adapt_delta <- 0.99
n_max_treedepth <- 20
## the max number of days after exposure to estimate
T_max <- 21
exposed_n <- 686
exposed_pos <- 77

source("R/utils.R")

## read in raw data
raw_data <- read_csv("data/antibody-test-data.csv") %>% 
    filter(grepl("RT_PCR", test),
           study != "Danis_no_4")

pcr_dat <- raw_data %>% 
    ## add non-quantified positives to other positives for Danis et al.
    mutate(n_adj=n+nqp,
           test_pos_adj=test_pos+nqp) %>% 
    ## remove estimates without observations
    filter(n_adj > 0,
           ## days needs to be above -5
           day > -5,
           ## only use the nasal swabs from Kujawski, not throat swabs
           !(study == "Kujawski" & test == "RT_PCR_oro")) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor() %>% as.numeric(),
           pct_pos=test_pos_adj/n_adj)

## make the data for a 3-day incubation period
pcr_dat3 <- raw_data %>% 
    mutate(n_adj=n+nqp,
           test_pos_adj=test_pos+nqp) %>% 
    filter(n_adj > 0,
           day > -3,
           !(study == "Kujawski" & test == "RT_PCR_oro")) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor() %>% as.numeric(),
           pct_pos=test_pos_adj/n_adj)

## make the data for a 7-day incubation period
pcr_dat7 <- raw_data %>% 
    mutate(n_adj=n+nqp,
           test_pos_adj=test_pos+nqp) %>% 
    filter(n_adj > 0,
           day > -7,
           !(study == "Kujawski" & test == "RT_PCR_oro")) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor() %>% as.numeric(),
           pct_pos=test_pos_adj/n_adj)

## only nasal swabs
naso_dat <- pcr_dat %>% 
    filter(grepl("RT_PCR_naso", test)) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor()
           %>% as.numeric(),
           pct_pos=test_pos_adj/n_adj)

## only throat swabs
oro_dat <- pcr_dat %>% 
    filter(grepl("RT_PCR_oro", test)) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor()
           %>% as.numeric(),
           pct_pos=test_pos_adj/n_adj)

## data for sensitivity analysis, where inconclusives are classified as negative
kuj_neg_dat <- raw_data %>% 
    mutate(n_adj = n+inconclusive+nqp,
           test_pos_adj = test_pos) %>% 
    filter(n_adj > 0,
           day > -5,
           !(study == "Kujawski" & test == "RT_PCR_oro")) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor()
           %>% as.numeric(),
           pct_pos=test_pos_adj/n_adj)

## data for sensitivity analysis, where inconclusives are classified as positive
kuj_pos_dat <- raw_data %>% 
    mutate(n_adj = n+inconclusive+nqp,
           test_pos_adj = test_pos+inconclusive+nqp) %>% 
    filter(grepl("RT_PCR", test),
           n_adj > 0,
           day > -5,
           !(study == "Kujawski" & test == "RT_PCR_oro")) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor()
           %>% as.numeric(),
           pct_pos=test_pos_adj/n_adj)

## create orthogonal polynomials for days since exposure
day_poly <- poly(log(pcr_dat$day+5), degree=3)
day_poly3 <- poly(log(pcr_dat3$day+3), degree=3)
day_poly7 <- poly(log(pcr_dat7$day+7), degree=3)
day_poly_guo <- poly(log(pcr_dat$day_min+5), degree=3)
day_poly_kuj <- poly(log(kuj_neg_dat$day_min+5), degree=3)

poly_predict <- predict(day_poly, log(1:T_max))
poly_predict3 <- predict(day_poly3, log(1:T_max))
poly_predict7 <- predict(day_poly7, log(1:T_max))
poly_predict_guo <- predict(day_poly_guo, log(1:T_max))
poly_predict_kuj <- predict(day_poly_kuj, log(1:T_max))


rv_dat <- raw_data %>% 
    ## add non-quantified positives to other positives for Danis et al.
    mutate(n_adj=n+nqp,
           test_pos_adj=test_pos+nqp) %>% 
    ## remove estimates without observations
    filter(n_adj > 0,
           ## days needs to be above -5
           day >= 0,
           ## only use the nasal swabs from Kujawski, not throat swabs
           !(study == "Kujawski" & test == "RT_PCR_oro")) %>% 
    mutate(study_idx=paste(study, test, sep="_") %>% as.factor() %>% as.numeric(),
           pct_pos=test_pos_adj/n_adj)

```

This code complements the submission "Variation in False Negative Rate of RT-PCR Based SARS-CoV-2 Tests by Time Since Exposure"

## Methods

[Zhao et al. (2020)](https://academic.oup.com/cid/advance-article/doi/10.1093/cid/ciaa344/5812996), [Liu et al. (2020)](https://www.medrxiv.org/content/10.1101/2020.03.06.20031856v1), [Guo et al. (2020)](https://academic.oup.com/cid/article-abstract/doi/10.1093/cid/ciaa310/5810754), [Wölfel et al. (2020)](https://www.nature.com/articles/s41586-020-2196-x), [Danis et al. (2020)](https://academic.oup.com/cid/article/doi/10.1093/cid/ciaa424/5819060), [Kujawski et al. (2020)](https://www.medrxiv.org/content/10.1101/2020.03.09.20032896v1.full), and [Kim et al. (2020)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7131901/) each looked at the sensitivity of the RT-PCR by time since symptom onset.

```{r raw-figures}
raw_data %>% 
    mutate(test=ifelse(test=="RT_PCR_oro", "Oropharyngeal",
                       ifelse(test=="RT_PCR_naso", "Nasopharyngeal",
                              "Unclear"))) %>% 
    ggplot(aes(x=day, y=(test_pos+nqp)/(n+nqp), size=n+nqp, color=study)) +
    facet_grid(.~test) +
    geom_point(alpha=0.5) +
    scale_x_continuous("Days since symptom onset",
                       breaks=seq(-7,35,7)) +
    scale_y_continuous("Raw sensitivity of the RT-PCR test",
                       labels=scales::percent) +
    scale_size_continuous("Number of\nobservations") +
    scale_color_manual("Study",
                       values=cbbPalette[-1]) +
    guides(colour = guide_legend(override.aes = list(alpha = 1))) +
    theme_bw() +
    theme(axis.text = element_text(color="black"),
          legend.position = "bottom",
          legend.title = element_text(size=10))
```

The sensitivity rises to a peak 4 days after symptom onset then declines for the next couple of weeks.

If we know the risk of an individual, we can find the negative predictive value -- the probability that someone who tests negative is actually negative.
From [Bi et al.](https://www.medrxiv.org/content/10.1101/2020.03.03.20028423v3), we know that about 15% (77/517) household contacts later tested positive for covid-19.

We use logistic regression for the sensitivity of the RT-PCR with a cubic polynomial for the log of time since exposure and use that, along with the probability of infection given exposure, to estimate the negative predictive value of the RT-PCR.
We use estimates of the incubation period from [Lauer, Grantz, et al. (2020)](https://annals.org/aim/fullarticle/2762808/incubation-period-coronavirus-disease-2019-covid-19-from-publicly-reported).
From this, we can find the probability of having a covid-19 infection despite being RT-PCR negative.

We use Stan for this analysis.

## Results

```{r stan-model, cache=T, include=F}
## fit a model to find the overall seroincidence across all observations
npv_onset_model <- stan_model("Stan/npv-fixed-onset.stan")
```


```{r npv-fixed-onset, cache=T, eval=T, include=F}
## run main analysis
main_analysis <- make_analysis_data(stan_model=npv_onset_model,
                                    dat=pcr_dat,
                                    T_max=T_max,
                                    poly_est=as.matrix(day_poly),
                                    poly_pred=poly_predict,
                                    exposed_n=exposed_n,
                                    exposed_pos=exposed_pos,
                                    spec=1,
                                    iter=n_iter,
                                    warmup=n_warmup,
                                    control=list(adapt_delta=p_adapt_delta,
                                                 max_treedepth=n_max_treedepth),
                                    save_warmup=F,
                                    save_stan=F)
```

```{r main-figure}
## check likelihood
# main_analysis$stan_ll

## false negative figure
fnr_fig <- ggplot(data=main_analysis$plot_dat, aes(x=days_since_exposure)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_errorbar(aes(ymin=fnr_lb, ymax=fnr_ub), color="gray30") +
    geom_point(aes(y=fnr_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability RT-PCR negative,\ngiven covid-19 positive",
                       limits=c(0,1)) +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())

## false omission figure
for_fig <- ggplot(data=main_analysis$plot_dat, aes(x=days_since_exposure)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_errorbar(aes(ymax=for_lb, ymin=for_ub), color="gray30") +
    geom_point(aes(y=for_med)) +
    scale_x_continuous("Days since exposure",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Post-test probability,\ngiven RT-PCR negative",
                       limits=c(0, 0.15),
                       breaks=seq(0,0.15, 0.05)) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))

gridExtra::grid.arrange(fnr_fig, for_fig)
```

```{r main-table}
## make analysis table
main_analysis$plot_dat %>% 
    rename(day=days_since_exposure) %>% 
    mutate_at(vars(-day), function(x) round(100*x,1)) %>% 
    knitr::kable()
```


With no data on RT-PCR to time prior to symptom onset, the estimates of sensitivity at or below day five are low with large credible intervals.
Due to the decline in sensitivity over time, the RT-PCR test is best deployed about a week after exposure.
A day or two after exposure (3 or 4 days prior to symptoms), the test may have no utility at all, and thus the probability of having been infected is would be the same with or without an RT-PCR, in our case about 11%.
Seven to nine days after exposure (roughly 2 to 4 days after symptom onset), the negative predictive value is around 95%, meaning there is about a 5% chance of actually being covid-19 positive despite testing negative.

```{r sub-est, cache=T, eval=F}
## separate models for subset analyses of nasal and throat swabs
## there weren't enough observations for the throat swab model to converge
naso_est <- make_analysis_data(stan_model=npv_onset_model,
                                    dat=naso_dat,
                                    T_max=T_max,
                                    poly_est=as.matrix(day_poly),
                                    poly_pred=poly_predict,
                                    exposed_n=exposed_n,
                                    exposed_pos=exposed_pos,
                                    spec=1,
                                    iter=n_iter,
                                    warmup=n_warmup,
                                    control=list(adapt_delta=p_adapt_delta,
                                                 max_treedepth=n_max_treedepth),
                                    save_warmup=F,
                                    save_stan=F)

oro_est <- make_analysis_data(stan_model=npv_onset_model,
                              dat=oro_dat,
                              T_max=T_max,
                              poly_est=as.matrix(day_poly),
                              poly_pred=poly_predict,
                              exposed_n=exposed_n,
                              exposed_pos=exposed_pos,
                              spec=1,
                              iter=n_iter,
                              warmup=n_warmup,
                              control=list(adapt_delta=p_adapt_delta,
                                           max_treedepth=n_max_treedepth),
                              save_warmup=F,
                              save_stan=F)
```

```{r plot-sub, eval=F}
bind_rows(naso_est$plot_dat %>% mutate(test="naso"),
          oro_est$plot_dat %>% mutate(test="oro"),
          main_analysis$plot_dat %>% mutate(test="original")) %>% 
    ggplot(aes(x=days_since_exposure, fill=as.factor(test),
               color=as.factor(test))) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_ribbon(aes(ymin=for_lb, ymax=for_ub), alpha=0.5) +
    geom_line(aes(y=for_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Post-test probability,\ngiven RT-PCR negative",
                       limits=c(0,0.15)) +
    scale_color_manual("Subset",
                       values=cbbPalette[c(2,1)]) +
    scale_fill_manual("Subset",
                      values=cbbPalette[c(2,1)]) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```

## Sensitivity analyses

### Specificity

What if the specificity of the test is less than 100%?
To test this we fit the same model except with a specificity of 90% and compared the results to the original (the sensitivity remained the same).

```{r sens-analysis-spec, cache=T, eval=T}
spec_est <- make_analysis_data(stan_model=npv_onset_model,
                               dat=pcr_dat,
                               T_max=T_max,
                               poly_est=as.matrix(day_poly),
                               poly_pred=poly_predict,
                               exposed_n=exposed_n,
                               exposed_pos=exposed_pos,
                               spec=0.9,
                               iter=n_iter,
                               warmup=n_warmup,
                               control=list(adapt_delta=p_adapt_delta,
                                            max_treedepth=n_max_treedepth),
                               save_warmup=F,
                               save_stan=F)
```

```{r plot-spec}
spec_est$plot_dat %>% 
    mutate(spec="90%") %>% 
    bind_rows(main_analysis$plot_dat %>% mutate(spec="100%")) %>% 
    ggplot(aes(x=days_since_exposure, color=spec, fill=spec)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_ribbon(aes(ymax=for_lb, ymin=for_ub), alpha=0.5) +
    # geom_errorbar(aes(ymax=for_lb, ymin=for_ub), alpha=0.5) +
    geom_line(aes(y=for_med)) +
    # geom_point(aes(y=for_med)) +
    scale_x_continuous("Days since exposure",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Post-test probability,\ngiven RT-PCR negative",
                       limits=c(0, 0.16),
                       breaks=seq(0,0.15, 0.05)) +
    scale_color_manual("Specificity",
                       values=cbbPalette) +
    scale_fill_manual("Specificity",
                      values=cbbPalette) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```

```{r spec-table}
spec_est$plot_dat %>% 
    select(day=days_since_exposure, for_90=for_med) %>% 
    left_join(main_analysis$plot_dat %>% 
                  select(day=days_since_exposure, for_med)) %>% 
    mutate(rr=for_90/for_med,
           abs_diff=for_90-for_med) %>% 
    mutate_at(vars(-day), function(x) round(100*x,1)) %>% 
    knitr::kable()
```

The shape of the curve for 90% specificity is similar to that of 100%, though slightly elevated.
The best time to test is still 2-4 days post-symptom onset.

\pagebreak

### Different pre-test probabilities

We tested pre-test probabilities of half, twice, and four times that of Bi et al. (2020).

```{r ar-est, cache=T, eval=T}
half_est <- make_analysis_data(stan_model=npv_onset_model,
                               dat=pcr_dat,
                               T_max=T_max,
                               poly_est=as.matrix(day_poly),
                               poly_pred=poly_predict,
                               exposed_n=exposed_n,
                               exposed_pos=round(exposed_pos/2),
                               spec=1,
                               iter=n_iter,
                               warmup=n_warmup,
                               control=list(adapt_delta=p_adapt_delta,
                                            max_treedepth=n_max_treedepth),
                               save_warmup=F,
                               save_stan=F)

two_est <- make_analysis_data(stan_model=npv_onset_model,
                              dat=pcr_dat,
                              T_max=T_max,
                              poly_est=as.matrix(day_poly),
                              poly_pred=poly_predict,
                              exposed_n=exposed_n,
                              exposed_pos=exposed_pos*2,
                              spec=1,
                              iter=n_iter,
                              warmup=n_warmup,
                              control=list(adapt_delta=p_adapt_delta,
                                           max_treedepth=n_max_treedepth),
                              save_warmup=F,
                              save_stan=F)

four_est <- make_analysis_data(stan_model=npv_onset_model,
                               dat=pcr_dat,
                               T_max=T_max,
                               poly_est=as.matrix(day_poly),
                               poly_pred=poly_predict,
                               exposed_n=exposed_n,
                               exposed_pos=exposed_pos*4,
                               spec=1,
                               iter=n_iter,
                               warmup=n_warmup,
                               control=list(adapt_delta=p_adapt_delta,
                                            max_treedepth=n_max_treedepth),
                               save_warmup=F,
                               save_stan=F)
```

```{r plot-ar}
bind_rows(half_est$plot_dat %>% mutate(ar_idx="half"),
          two_est$plot_dat %>% mutate(ar_idx="2x"),
          four_est$plot_dat %>% mutate(ar_idx="4x"),
          main_analysis$plot_dat %>% mutate(ar_idx="Bi")) %>% 
    ggplot(aes(x=days_since_exposure, fill=as.factor(ar_idx),
               color=as.factor(ar_idx))) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    # geom_ribbon(aes(ymin=for_lb, ymax=for_ub), alpha=0.3) +
    geom_line(aes(y=for_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Post-test probability,\ngiven RT-PCR negative") +
    scale_color_manual("Pre-test\nprobability",
                       values=cbbPalette[c(2,3,1,4)],
                       breaks=c("half", "Bi", "2x", "4x"),
                       labels=c("5.5%", "11%", "22%", "44%")) +
    scale_fill_manual("Pre-test\nprobability",
                      values=cbbPalette[c(2,3,1,4)],
                      breaks=c("half", "Bi", "2x", "4x"),
                      labels=c("5.5%", "11%", "22%", "44%")) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```

```{r ar-table}
bind_rows(half_est$plot_dat %>% mutate(ar_idx="half"),
          two_est$plot_dat %>% mutate(ar_idx="2x"),
          four_est$plot_dat %>% mutate(ar_idx="4x")) %>% 
    select(day=days_since_exposure, ar_idx, for_med, for_lb, for_ub) %>% 
    filter(day<10) %>% 
    mutate_at(vars(starts_with("for")), function(x) round(100*x,1)) %>% 
    knitr::kable()
```

As the pre-test probability increases, the shape looks more like the false negative rate seen above in Figure 2.
As the pre-test probability decreases, the probabilities approach 0 and thus the entire curve flattens out.
Regardless, the best time to test is still 2-4 days after symptom onset.

\pagebreak

### Different incubation period lengths

We originally assumed a 5-day incubation period, what if that was 3 or 7 days instead?

```{r inc-est, cache=T, eval=T}
three_day_est <- make_analysis_data(stan_model=npv_onset_model,
                                    dat=pcr_dat3,
                                    T_max=T_max,
                                    poly_est=as.matrix(day_poly3),
                                    poly_pred=poly_predict3,
                                    exposed_n=exposed_n,
                                    exposed_pos=exposed_pos,
                                    spec=1,
                                    iter=n_iter,
                                    warmup=n_warmup,
                                    control=list(adapt_delta=p_adapt_delta,
                                                 max_treedepth=n_max_treedepth),
                                    save_warmup=F,
                                    save_stan=F)

seven_day_est <- make_analysis_data(stan_model=npv_onset_model,
                                    dat=pcr_dat7,
                                    T_max=T_max,
                                    poly_est=as.matrix(day_poly7),
                                    poly_pred=poly_predict7,
                                    exposed_n=exposed_n,
                                    exposed_pos=exposed_pos,
                                    spec=1,
                                    iter=n_iter,
                                    warmup=n_warmup,
                                    control=list(adapt_delta=p_adapt_delta,
                                                 max_treedepth=n_max_treedepth),
                                    save_warmup=F,
                                    save_stan=F)
```

```{r plot-inc}
bind_rows(three_day_est$plot_dat %>% mutate(inc_period="3d"),
          seven_day_est$plot_dat %>% mutate(inc_period="7d"),
          main_analysis$plot_dat %>% mutate(inc_period="5d")) %>% 
    ggplot(aes(x=days_since_exposure, fill=as.factor(inc_period),
               color=as.factor(inc_period))) +
    geom_vline(aes(xintercept=3), linetype="dashed", color=cbbPalette[2]) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_vline(aes(xintercept=7), linetype="dashed", color=cbbPalette[3]) +
    # geom_ribbon(aes(ymin=for_lb, ymax=for_ub), alpha=0.1) +
    geom_line(aes(y=for_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Post-test probability,\ngiven RT-PCR negative",
                       limits=c(0,0.15)) +
    scale_color_manual("Incubation period",
                       values=cbbPalette[c(2,1,3)]) +
    scale_fill_manual("Incubation period",
                      values=cbbPalette[c(2,1,3)]) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```

```{r inc-table}
bind_rows(three_day_est$plot_dat %>% mutate(inc_period="3d"),
          seven_day_est$plot_dat %>% mutate(inc_period="7d")) %>%
    select(day=days_since_exposure, inc_period, for_med, for_lb, for_ub) %>% 
    filter(day<14) %>% 
    mutate_at(vars(starts_with("for")), function(x) round(100*x,1)) %>% 
    knitr::kable()
```

Changing the length of the incubation period changes the progression of the false omission rate (post-test probability given test negative).
Since the sensitivity is calibrated with respect to the time of symptom onset, an earlier onset time leads to a quicker drop in false omission rate and a later onset time leads to a slower drop.

\pagebreak

### Shift ambiguous days one day earlier

The timing of the days since symptom onset are ambiguous in Guo et al. and Kim et al., where day 1 may mean one day since symptom onset or the first day of symptoms.

```{r guo-est, cache=T, eval=T}
guo_est <- make_analysis_data(stan_model=npv_onset_model,
                              dat=pcr_dat,
                              T_max=T_max,
                              poly_est=as.matrix(day_poly_guo),
                              poly_pred=poly_predict_guo,
                              exposed_n=exposed_n,
                              exposed_pos=exposed_pos,
                              spec=1,
                              iter=n_iter,
                              warmup=n_warmup,
                              control=list(adapt_delta=p_adapt_delta,
                                           max_treedepth=n_max_treedepth),
                              save_warmup=F,
                              save_stan=F)
```

```{r plot-guo, eval=T}
# guo_est$stan_ll
for_guo <- bind_rows(guo_est$plot_dat %>% mutate(guo_day="0d"),
                     main_analysis$plot_dat %>% mutate(guo_day="1d")) %>% 
    ggplot(aes(x=days_since_exposure, fill=as.factor(guo_day),
               color=as.factor(guo_day))) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_ribbon(aes(ymin=for_lb, ymax=for_ub), alpha=0.5) +
    geom_line(aes(y=for_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Post-test probability,\ngiven RT-PCR negative",
                       limits=c(0,0.15)) +
    scale_color_manual("Guo first day",
                       values=cbbPalette[c(2,1)]) +
    scale_fill_manual("Guo first day",
                      values=cbbPalette[c(2,1)]) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))

fnr_guo <- bind_rows(guo_est$plot_dat %>% mutate(guo_day="0d"),
                     main_analysis$plot_dat %>% mutate(guo_day="1d")) %>% 
    ggplot(aes(x=days_since_exposure, fill=as.factor(guo_day),
               color=as.factor(guo_day))) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_ribbon(aes(ymin=fnr_lb, ymax=fnr_ub), alpha=0.5) +
    geom_line(aes(y=fnr_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability RT-PCR negative,\ngiven covid-19 positive",
                       limits=c(0,1)) +
    scale_color_manual("Guo first day",
                       values=cbbPalette[c(2,1)]) +
    scale_fill_manual("Guo first day",
                      values=cbbPalette[c(2,1)]) +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())

gridExtra::grid.arrange(fnr_guo, for_guo)

```

Shifting the timing of ambiguous days one day earlier improves the median sensitivity of the early tests, such that the best days for testing would shift from 2-4 days post-symptom onset to 1-3 days post-sypmtom onset.
The log-likelihoods for each model are nearly identical with the shifted days being minutely better.
The overlap in credible intervals indicates that the estimates are roughly equivalent.

\pagebreak

### Inconclusive tests as negatives or positives

Kujawski et al. and Danis et al. reported `r sum(pcr_dat$inconclusive)` inconclusive tests combined in their data.
In our main analysis, we omit these swabs, but they could count as negative or positive tests instead.

```{r kuj-est, cache=T, eval=T}
kuj_neg <- make_analysis_data(stan_model=npv_onset_model,
                              dat=kuj_neg_dat,
                              T_max=T_max,
                              poly_est=as.matrix(day_poly_kuj),
                              poly_pred=poly_predict_kuj,
                              exposed_n=exposed_n,
                              exposed_pos=exposed_pos,
                              spec=1,
                              iter=n_iter,
                              warmup=n_warmup,
                              control=list(adapt_delta=p_adapt_delta,
                                           max_treedepth=n_max_treedepth),
                              save_warmup=F,
                              save_stan=F)

kuj_pos <- make_analysis_data(stan_model=npv_onset_model,
                              dat=kuj_pos_dat,
                              T_max=T_max,
                              poly_est=as.matrix(day_poly_kuj),
                              poly_pred=poly_predict_kuj,
                              exposed_n=exposed_n,
                              exposed_pos=exposed_pos,
                              spec=1,
                              iter=n_iter,
                              warmup=n_warmup,
                              control=list(adapt_delta=p_adapt_delta,
                                           max_treedepth=n_max_treedepth),
                              save_warmup=F,
                              save_stan=F)
```

```{r plot-kuj, eval=T}
# print("Likelihood for inconclusives as negatives")
# kuj_neg$stan_ll
# print("Likelihood for inconclusives as positives")
# kuj_pos$stan_ll
# loo::loo_compare(kuj_neg$stan_ll, kuj_pos$stan_ll)
bind_rows(kuj_neg$plot_dat %>% mutate(kuj="negative"),
          kuj_pos$plot_dat %>% mutate(kuj="positive"),
          main_analysis$plot_dat %>% mutate(kuj="omitted")) %>% 
    ggplot(aes(x=days_since_exposure, fill=as.factor(kuj),
               color=as.factor(kuj))) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    # geom_ribbon(aes(ymin=for_lb, ymax=for_ub), alpha=0.5) +
    geom_line(aes(y=for_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Post-test probability,\ngiven RT-PCR negative",
                       limits=c(0,0.15)) +
    scale_color_manual("Inconclusives as",
                       values=cbbPalette[c(2,1,3)]) +
    scale_fill_manual("Inconclusives as",
                      values=cbbPalette[c(2,1,3)]) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```

The results are barely changed by including inconclusive tests as negatives or positives instead of being omitted.

### Leave one study out

To see whether any study has undue influence over our results, we ran a leave-one-study out analysis to see if the results changed


```{r loso-est, cache=T, eval=T}
n_studies <- unique(pcr_dat$study)
loso_dat <- c()
for(i in n_studies){
    tmp_loso <- filter(pcr_dat, study != i)
    tmp_poly <- poly(log(tmp_loso$day+5), degree=3)
    tmp_poly_pred <- predict(tmp_poly, log(1:T_max))
    tmp_est <- make_analysis_data(stan_model=npv_onset_model,
                                  dat=tmp_loso,
                                  T_max=T_max,
                                  poly_est=as.matrix(tmp_poly),
                                  poly_pred=tmp_poly_pred,
                                  exposed_n=exposed_n,
                                  exposed_pos=exposed_pos,
                                  spec=1,
                                  iter=n_iter,
                                  warmup=n_warmup,
                                  control=list(adapt_delta=p_adapt_delta,
                                               max_treedepth=n_max_treedepth),
                                  save_warmup=F,
                                  save_stan=F)
    loso_dat <- bind_rows(loso_dat,
                          tmp_est$plot_dat %>% mutate(left_out=i))
}
```

```{r plot-loso, eval=T}
bind_rows(loso_dat,
    main_analysis$plot_dat %>% mutate(left_out="none")) %>% 
    mutate(left_out=relevel(factor(left_out), "none")) %>% 
    ggplot(aes(x=days_since_exposure, fill=left_out, color=left_out)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    # geom_ribbon(aes(ymin=for_lb, ymax=for_ub), alpha=0.5) +
    geom_line(aes(y=for_med), alpha=0.5) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Post-test probability,\ngiven RT-PCR negative",
                       limits=c(0,0.15)) +
    scale_color_manual("Left out",
                       values=cbbPalette) +
    scale_fill_manual("Left out",
                      values=cbbPalette) +
    theme_bw() +
    guides(colour = guide_legend(override.aes = list(alpha = 1))) +
    theme(axis.text=element_text(color="black"))
```
