---
title: "Probability of covid-19 infection given RT-PCR negative"
author: "Stephen A Lauer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
knitr::opts_knit$set(root.dir = here::here())

options(mc.cores=4,
        scipen=999)
```

```{r library}
library(tidyverse)
library(rstan)

n_iter <- 1e4
n_warmup <- 1e4-5e3
p_adapt_delta <- 0.99
n_max_treedepth <- 20
T_max <- 21

test_dat <- read_csv("data/antibody-test-data.csv") %>% 
    mutate(study_idx=study %>% as.factor() %>% as.numeric())
```

In this paper, we will try to determine the probability that someone has covid-19 given a negative RT-PCR test.

### Methods

[Zhao et al. (2020)](https://academic.oup.com/cid/advance-article/doi/10.1093/cid/ciaa344/5812996) and [Liu et al. (2020)](https://www.medrxiv.org/content/10.1101/2020.03.06.20031856v1) looked at the sensitivity of the RT-PCR (and ELISAs) by time since symptom onset.

```{r raw-figures}
ggplot(data=test_dat, aes(x=day, y=pct_pos, size=n, color=study)) +
    geom_point() +
    scale_x_continuous("Days since symptom onset",
                       breaks=seq(0,21,7)) +
    scale_y_continuous("Sensitivity of the RT-PCR test") +
    scale_size_continuous("Number of\nobservations") +
    theme_bw()
```

The sensitivity rises to a peak 4 days after symptom onset then declines for the next couple of weeks.

If we know the risk of an individual, we can find the negative predictive value -- the probability that someone who tests negative is actually negative.
From [Bi et al.](https://www.medrxiv.org/content/10.1101/2020.03.03.20028423v3), we know that about 15% (77/517) household contacts later tested positive for covid-19.

We use logistic regression for the sensitivity of the RT-PCR with a cubic polynomial for the log of time since exposure and use that, along with the probability of infection given exposure, to estimate the negative predictive value of the RT-PCR.
We use estimates of the incubation period from [Lauer, Grantz, et al. (2020)](https://annals.org/aim/fullarticle/2762808/incubation-period-coronavirus-disease-2019-covid-19-from-publicly-reported).
From this, we can find the probability of having a covid-19 infection despite being RT-PCR negative.

We use Stan for this analysis.

### Results

```{r fixed-onset-stan-b, cache=T, eval=T}
## fit a model to find the overall seroincidence across all observations
npv_onset_model <- stan_model("Stan/npv-fixed-onset.stan")

# test_dat2 <- test_dat %>% 
#     group_by(day) %>% 
#     summarize(test_pos=sum(test_pos),
#               n=sum(n))

npv_est <- sampling(npv_onset_model,
                    data=list(N=nrow(test_dat),
                              J=max(test_dat$study_idx),
                              T_max=T_max,
                              test_n=test_dat$n,
                              test_pos=test_dat$test_pos,
                              study_idx=test_dat$study_idx,
                              t_symp_test=test_dat$day,
                              exposed_n=686,
                              exposed_pos=77,
                              t_exp_symp=5
                    ),
                    iter=n_iter,
                    warmup=n_warmup,
                    control=list(adapt_delta=p_adapt_delta,
                                 max_treedepth=n_max_treedepth),
                    save_warmup=F)

loo::extract_log_lik(npv_est) %>% loo::loo()

rm(npv_onset_model)
```

```{r all-conditionals}
# loo::extract_log_lik(npv_est) %>% loo::loo()
## sensitivity (sens) of PCR: P(PCR+ | covid+)
## false negative rate (fnr) of PCR: P(PCR- | covid+) = 1 - sens
sens <- extract(npv_est, pars="sens")[[1]]

## negative predictive value (npv) of PCR: P(covid- | PCR-)
## false omission rate (FOR) of PCR: P(covid+ | PCR-) = 1 - npv
npv <- extract(npv_est, pars="npv")[[1]]

## incubation time
# inc_time <- extract(npv_est, pars="t_exp_symp")[[1]]
inc_time <- 5

## attack rate: P(covid+)
## P(covid-) = 1 - attack_rate
attack_rate <- extract(npv_est, pars="attack_rate")[[1]] %>% as.vector()
```


```{r plot-dat}
plot_dat <- as_tibble(sens) %>%
    gather("days", "sens") %>%
    mutate(days_since_exposure=gsub(pattern="V", "", days) %>% as.numeric) %>%
    bind_cols(as_tibble(npv) %>%
                  gather("days", "npv") %>%
                  mutate(ar=rep(attack_rate, T_max)) %>% 
                  select(-days)) %>% 
    group_by(days_since_exposure) %>%
    summarise(fnr_med=median(1-sens),
              fnr_lb=quantile(1-sens,probs=.025),
              fnr_ub=quantile(1-sens,probs=.975),
              for_med=median(1-npv),
              for_lb=quantile(1-npv,probs=.025),
              for_ub=quantile(1-npv,probs=.975),
              rr_med=median(1-(1-npv)/ar),
              rr_lb=quantile(1-(1-npv)/ar,probs=.025),
              rr_ub=quantile(1-(1-npv)/ar,probs=.975),
              abs_med=median(ar-(1-npv)),
              abs_lb=quantile(ar-(1-npv),probs=.025),
              abs_ub=quantile(ar-(1-npv),probs=.975))

paste0("Median (95% CI) of attack rate: ", median(100*attack_rate) %>% round(), "% (",
       quantile(100*attack_rate, probs=.025) %>% round(), ", ",
       quantile(100*attack_rate, probs=.975) %>% round(), "%)")
knitr::kable(plot_dat %>%
                 rename(Days=days_since_exposure) %>% 
                 mutate_at(vars(2:ncol(plot_dat)), function(x) paste0(round(100*x), "%")),
             digits=0)
```

```{r fnr-figure}
fnr_fig <- ggplot(data=plot_dat, aes(x=days_since_exposure)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_errorbar(aes(ymin=fnr_lb, ymax=fnr_ub), color="gray30") +
    geom_point(aes(y=fnr_med)) +
    scale_x_continuous("",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability RT-PCR negative,\ngiven covid-19 positive",
                       limits=c(0,1)) +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())
```

With no data on RT-PCR to time prior to symptom onset, the estimates of sensitivity at or below day five are low with large credible intervals.

```{r npv-figure}
for_fig <- ggplot(data=plot_dat, aes(x=days_since_exposure)) +
    geom_vline(aes(xintercept=5), linetype="dashed") +
    geom_errorbar(aes(ymax=for_lb, ymin=for_ub), color="gray30") +
    geom_point(aes(y=for_med)) +
    scale_x_continuous("Days since exposure",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_continuous("Probability infected,\ngiven RT-PCR negative",
                       limits=c(0, 0.15),
                       breaks=seq(0,0.15, 0.05)) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))

gridExtra::grid.arrange(fnr_fig, for_fig)

ggplot(plot_dat, aes(x=days_since_exposure)) +
    geom_hline(aes(yintercept=1), linetype="dashed") +
    geom_errorbar(aes(ymax=1-rr_lb, ymin=1-rr_ub), color="gray30") +
    geom_point(aes(y=1-rr_med)) +
    scale_x_continuous("Days since exposure",
                       breaks=seq(0, 21, 7),
                       limits=c(0,21.5)) +
    scale_y_log10("Relative risk of having a negative RT-PCR test",
                  breaks=2^c(-3:1),
                  labels=c("1/8", "1/4", "1/2", "1", "2")) +
    coord_cartesian(ylim=c(2^-3, 2^1)) +
    theme_bw() +
    theme(axis.text=element_text(color="black"))
```


Due to the decline in sensitivity over time, the RT-PCR test is best deployed about a week after exposure.
A day or two after exposure (3 or 4 days prior to symptoms), the test may have no utility at all, and thus the probability of having been infected is would be the same with or without an RT-PCR, in our case about 15%.
Seven to nine days after exposure (roughly 2 to 4 days after symptom onset), the negative predictive value is around 95%, meaning there is about a 5% chance of actually being covid-19 positive despite testing negative.
